---
title: "coding_exercise"
author: "Katie Frazer"
date: "2024-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("setup.R")
```

For these exercises, we will be using the **knz_bison** dataset from the **lterdatasampler** package.

```{r}
data("knz_bison")

knz_bison
```

Create a new column in the bison dataset that calculates the individual's age at the time of observation.

```{r}
# only use year info as month and day info unavailable for animal's birthdate
# animal_yob ~ a number denoting the year animal was born
# rec_year ~ a number denoting the year of observation

knz_bison %>%
  mutate(animal_age_obs = rec_year-animal_yob)
```

Write a function that will first filter the bison dataset based on any user-specified weight threshold (returning the filtered dataset).

```{r}
# animal_weight ~ a number denoting bison weight in pounds
# filtering as if user wants anything less than or equal to provided weight

bison_by_weight <- 
  function(x){ 
    knz_bison %>%
    filter(animal_weight <= x)
  }

bison_by_weight(300)
```

Create a boxplot showing weight distribution between male and female bison, and then conduct a t-test to see if weight is significantly different between males and females.

```{r}
# animal_weight ~ a number denoting bison weight in pounds
# animal_sex ~ a character denoting the sex of the bison
  # M = male, F = female

knz_bison %>%
  ggplot(aes(x = animal_weight)) +
  geom_boxplot(aes(fill = animal_sex))
```

```{r}
# running a t-test

# testing for equal variances
knz_bison %>%
  levene_test(animal_weight ~ animal_sex)
  # variances are NOT equal

# doing the t-test 
knz_bison %>%
  t_test(animal_weight ~ animal_sex, var.equal = FALSE, detailed = TRUE)
  # bison weight is significantly higher in female bison over male mison
```
